---
title: "mortality"
output: html_document
date: "2024-03-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# install.packages("WDI")
library(WDI)
library(tidyverse)
# WDIsearch("infant") # yields the indicator: SP.DYN.IMRT.IN
# WDIsearch("GDP") # yields the indicator: NY.GDP.PCAP.CD
InfantMort = WDI(,"SP.DYN.IMRT.IN",start = 2019, end = 2019)
GDPs = WDI(,"NY.GDP.PCAP.CD",start = 2019, end = 2019)
InfantVGDP = inner_join(GDPs,InfantMort, "country")
InfantVGDP = InfantVGDP %>% select("country","NY.GDP.PCAP.CD","SP.DYN.IMRT.IN")
colnames(InfantVGDP) = c("country","GDP","InfantMort")
# InfantVGDP
```

# GDP of the United States of America
```{r}
usa <- grep("United States", InfantVGDP$country)
usa
InfantVGDP[usa,]
```
# Log USA GDP, Infant Mortality Rate
```{r}
logGDP <- log(InfantVGDP[usa,]$GDP)
logInfantMort <- log(InfantVGDP[usa,]$InfantMort)
```


# ANOVA
```{r}
fit <- lm(InfantMort~GDP, data=InfantVGDP)
anova(fit)
```

# Fit Summary
```{r}
summary(fit)
```

# Log-Linear transform
```{r}
InfantVGDP.logInfantMort <- log(InfantVGDP$InfantMort)
fit1 <- lm(InfantVGDP.logInfantMort~GDP, data=InfantVGDP)
summary(fit1)
```
# Log-Linear plot
```{r}
plot(fit1, main="Log-Linear Residuals")
```
# Log-Linear Residual Plot
```{r}
plot(fit1$residuals, main="Log-Linear Residual Plot", xlab="Predicted Values", ylab="Residuals")
abline(h=0, col="black")
```
# Log-Linear Regression Model
```{r}
plot(InfantVGDP$GDP, InfantVGDP.logInfantMort, main="Log-Linear Regression Model: Log(Infant Mortality) ~ GDP", xlab="GDP", ylab="Log(Infant Mortality Rate)")
abline(fit1, col="red")
```

# Linear-Log transform
```{r}
InfantVGDP.logGDP <- log(InfantVGDP$GDP)
fit2 <- lm(InfantMort~InfantVGDP.logGDP, data=InfantVGDP)
summary(fit2)
```
# Linear-Log Residual plot
```{r}
plot(fit2, main="Linear-Log Residual Plot")
```
# Linear-Log Residual plot
```{r}
plot(fit2$residuals, main="Linear-Log Residual Plot", xlab="Predicted Values", ylab="Residuals")
abline(h=0, col="black")
```
# Linear-Log regression model
```{r}
plot(InfantVGDP.logGDP, InfantVGDP$InfantMort, main="Linear-Log Model: Infant Mortality ~ Log(GDP)", xlab="Log(GDP)", ylab="Infant Mortality Rate")
abline(fit2, col="red")
```



# Log-Log transform
```{r}
fit3 <- lm(InfantVGDP.logInfantMort~InfantVGDP.logGDP, data=InfantVGDP)
summary(fit3)
```
# Log-Log Residual plot
```{r}
plot(fit3, main="Log-Log Residual Plot")
```
# Log-Log Residual plot
```{r}
plot(fit3$residuals, main="Log-Log Residual", xlab="Predicted Values", ylab="Residuals")
abline(h=0, col="black")
```
# Log-Log Regression Model
```{r}
plot(InfantVGDP.logGDP, InfantVGDP.logInfantMort, main="Log-Log Regression Model: Log(Infant Mortality) ~ Log(GDP)", xlab="Log(GDP)", ylab="Log(Infant Mortality Rate)")
abline(fit3, col="red")
points(logGDP, logInfantMort, col="blue", pch=19)
text(logGDP, logInfantMort, labels = paste("USA (", round(logGDP,2), ",", round(logInfantMort, 2), ")", sep = ""), pos = 4, cex=0.5, col="blue")
```
```{r}

#                  Estimate Std. Error t value Pr(>|t|)    
# (Intercept)        8.33288    0.22429   37.15   <2e-16 ***
# InfantVGDP.logGDP -0.65519    0.02544  -25.75   <2e-16 ***

# SLOPE
a <- 0.975
n <- nrow(InfantVGDP)
df <- n - 2
cv <- qt(a, df)
se_beta_0 <- 0.22429
se_beta_1 <- 0.02544

beta_1 <- fit3$coefficients[2]
beta_1_lower <- beta_1 - se_beta_1*cv
beta_1_upper <- beta_1 + se_beta_1*cv

2^beta_1_lower
2^beta_1_upper

# INTERCEPT

beta_0 <- fit3$coefficients[1]
beta_0_lower <- beta_0 - se_beta_0*cv
beta_0_upper <- beta_0 + se_beta_0*cv

2^beta_0_lower
2^beta_0_upper
```



