---
title: "model"
author: "ddxbugs"
date: "2024-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

sessionInfo()
getwd()
setwd("../autism")
```

```{r}
df <- read.csv("Autism Data Prob 29.csv")
# str(df)
# View(df)
```

# Scatter plot: Autism Prevalence ~ Years(1992-2000)
```{r}
plot(df, main="Scatter plot: Autism Prevalence ~ Years(1999-2000)")
```

# Simple linear regression model
```{r}
fit <- lm(Prevalence~Year, data=df)
fit
# str(fit)
```

# Summary fit
```{r}
summary(fit)
```

# ANOVA fit
```{r}
anova(fit)
```
# Residual plots
```{r}
plot(fit)
```
# Residual plot
```{r}
plot(fit$residuals, ylab="Residuals", main="Residual Plot")
abline(h=0, col="black")
```
# Histogram of residuals
```{r}
x_resids <- mean(fit$residuals)
s_resids <- sqrt(var(fit$residuals))
hist(fit$residuals, density=20, breaks=5, prob=TRUE, col="red",
     xlab="Residuals", ylim=c(0,1),
     main=paste("Histogram of Residuals ~ N(",round(x_resids,3),", ",round(s_resids,3),")"))
curve(dnorm(x, mean=x_resids, sd=s_resids), col="darkblue", lwd=2, add=TRUE, yaxt="n")
```
# Log-linear transformation
```{r}
df1 <- data.frame(Year=df$Year, logPrev=log(df$Prevalence))
# str(df1)
```

# Scatter plot: ln(Autism Prevalence) ~ Years(1992-2000)
```{r}
plot(df1, main="Scatter plot: ln(Autism Prevalence)~Years(1999-2000)", ylab="ln(Prevelance)")
```
# Log-linear regression model
```{r}
fit1 <- lm(logPrev~Year, data=df1)
fit1
# str(fit1)
```
# Summary fit
```{r}
summary(fit1)
```

# Log-linear residual plots
```{r}
plot(fit1)
```

# Fitted plot: Autism Prevalence ~ Years(1999-2000)
```{r}
newx1 <- df$Year
newx1 <- sort(newx1)

pred_x1 <- predict(fit, newdata=data.frame(Year=newx1), interval="confidence", type="response", level=0.95)
pred_y1 <- predict(fit, newdata=data.frame(Year=newx1), interval="prediction", type="response", level=0.95)

plot(df, xlim=c(1991, 2001), ylim=c(0, 20), xlab="Year", ylab="Prevelance", main="Fitted Plot: Prevelance ~ Years", cex=0.5)
abline(fit, col="red")
lines(newx1, pred_x1[,2], col="black", lty=2, lwd=1)
lines(newx1, pred_x1[,3], col="black", lty=2, lwd=1)
lines(newx1, pred_y1[,2], col="green", lty=2, lwd=2)
lines(newx1, pred_y1[,3], col="green", lty=2, lwd=2)
```

# Fitted plot: ln(Autism Prevalence)~Years(1999-2000)
```{r}
newx2 <- df1$Year
newx2 <- sort(newx2)

pred_x2 <- predict(fit1, newdata=data.frame(Year=newx2), interval="confidence", type="response", level=0.95)
pred_y2 <- predict(fit1, newdata=data.frame(Year=newx2), interval="prediction", type="response", level=0.95)

plot(df1, xlim=c(1991, 2001), ylim=c(0, 3), xlab="Year", ylab="ln(Prevelance)", main="Fitted Plot: ln(Prevelance) ~ Years", cex=0.5)
abline(fit1, col="red")
lines(newx2, pred_x2[,2], col="black", lty=2, lwd=1)
lines(newx2, pred_x2[,3], col="black", lty=2, lwd=1)
lines(newx2, pred_y2[,2], col="green", lty=2, lwd=2)
lines(newx2, pred_y2[,3], col="green", lty=2, lwd=2)
```
# Histogram of log-linear residuals
```{r}
x_resids2 <- mean(fit1$residuals)
s_resids2 <- sqrt(var(fit1$residuals))
hist(fit1$residuals, density=20, breaks=5, prob=TRUE, col="red",
     xlab="Residuals", ylim=c(0,50),
     main=paste("Histogram of Log-linear Residuals ~ N(", round(x_resids2,3), ",",round(s_resids2,3), ")"))
curve(dnorm(x, mean=x_resids2, sd=s_resids2), col="darkblue", lwd=2, add=TRUE, yaxt="n")
```
# Coefficients
```{r}
# Coefficients:
#              Estimate Std. Error t value Pr(>|t|)    
# (Intercept) -4.080e+02  4.953e+00  -82.38 3.94e-06 ***
# Year         2.054e-01  2.481e-03   82.79 3.88e-06 ***

a <- 0.975
n <- nrow(df1)
N <- n - 2
cv <- qt(a, N)

se_beta_0 <- 4.953
se_beta_1 <- 0.002481
```

# Beta_1
```{r}
# SLOPE
beta_1 <- fit1$coefficients[2]
beta_1_lower <- beta_1 - se_beta_1*cv
beta_1_upper <- beta_1 + se_beta_1*cv

beta_1_lower
beta_1_upper
```

# Beta_0
```{r}
# INTERCEPT
beta_0 <- fit1$coefficients[1]
beta_0_lower <- beta_0 - se_beta_0*cv
beta_0_upper <- beta_0 + se_beta_0*cv

beta_0_lower
beta_0_upper
```

# Test
```{r}
year <- 2000
pred_prev <- beta_0 + beta_1*year
exp(pred_prev)
```







